---
title: "Exploration of Metabolite Levels Across the Brains of Rats Treated with Varriable Doses of DMSO"
author: Zach Rabow and Hannah Houts
date: March 15, 2019
output: 
  ioslides_presentation:
    transition: slower
    incremental: true
#    css: styles.css

autosize: true
---

# Description of the Experiment

 - This is a description of the study, and how rats got injected with solvents. its a whole thing

#Data Tidying
## Description of the Data 

The data was came as an excel file that was not #Tidy at all. 

-data (treatment) was stored in column names, not as a vector
-this occured across 3 parts of the data:
  -the mean intensity
  -the p-value
  -the fold change
-vector names are full of "."s and spaces, which will cause problems with ggplot
  
## How we Tidyd the data
```{r load data & libraries, include = FALSE}
Untidy_DMSO_Data <- readr::read_csv("UntidyFYSDataSet.csv")

library("tidyverse")
library("ggplot2")
library("ggthemr")
library("devtools")
library("RColorBrewer")

options(scipen = 999)
```

```{r}
Untidy_DMSO_Data
```

## How we #Tidyd the data
Oof, those column names have to go
```{r}
colnames(Untidy_DMSO_Data)
```

## How we #Tidyd the data
```{r}
Untidy_DMSO_Data <- Untidy_DMSO_Data %>%
  rename(
          "Chemical_Name" = ChemicalName,
          "Mean_Intensity_Control" = mean.Control.PBS,
          "Mean_Intensity_02" = mean.0.2.mg.kg.DMSO,
          "Mean_Intensity_20" = mean.2.0.mg.kg.DMSO,
          "Mean_Intensity_40" = mean.4.0.mg.kg.DMSO,
          "Fold_Change_02" = FC_lowvsCTRL,
          "Fold_Change_20" = FC_medvsCTRL,
          "Fold_Change_40" = FC_highvsCTRL,
          "p_val_02" = post.hoc.Control.PBS.0.2.mg.kg.DMSO,
          "p_val_20" = post.hoc.Control.PBS.2.0.mg.kg.DMSO,
          "p_val_40" = post.hoc.Control.PBS.4.0.mg.kg.DMSO
          )
```

## How we #Tidyd the data
Much Better!
```{r}
colnames(Untidy_DMSO_Data)
```
Now we need to Gather some columns

## How we #Tidyd the data

We gathered intensities and made a data frame with <b> only </b> intensities
```{r, results='hide'}
Intensity_DMSO_Data<- Untidy_DMSO_Data %>% 
   #Gathers 4 columns 
   gather(Mean_Intensity_Control, 
         Mean_Intensity_02, 
         Mean_Intensity_20, 
         Mean_Intensity_40, 
         key="Treatment", value="Mean_Intensity")%>% 
  
   #deletes part of unwanted text in data
   separate(Treatment, 
            into= c("Delete", "Treatment"), 
            sep = "_Intensity_") %>%   
  
   # selects 3 "key" columns, treatment, and intensity
   select(Region, ID, Chemical_Name, Treatment, Mean_Intensity)
```

## How we #Tidyd the data

We gathered fold changes and made a data frame with <b> only </b> fold changes
```{r, results='hide'}
Fold_Change_DMSO_Data<- Untidy_DMSO_Data %>% 
 #Gathers 4 columns 
   gather(Fold_Change_02, 
           Fold_Change_20,
           Fold_Change_40,
          key="Treatment", value="Fold_Change_vs_Control") %>% 

 #deletes part of unwanted text in data
 separate(Treatment, 
            into= c("Delete", "Treatment"), 
            sep = "_Change_") %>%   
  
 # selects 3 "key" columns, treatment, and intensity
 select(Region, ID, Chemical_Name, Treatment, Fold_Change_vs_Control)
```

## How we #Tidyd the data

We gathered p-values and made a data frame with <b> only </b> p-values
```{r, results='hide'}
p_value_DMSO_Data<- Untidy_DMSO_Data %>% 
 #Gathers 3 columns 
 gather(p_val_02, 
        p_val_20, 
        p_val_40,
        key="Treatment", value="p_value_vs_Control") %>% 
  
 #deletes part of unwanted text in data
 separate(Treatment, 
            into= c("Delete", "Treatment"), 
            sep = "_val_") %>%   
 # selects 3 "key" columns, treatment, and intensity
 select(Region, ID, Chemical_Name, Treatment, p_value_vs_Control)

```
## How we #Tidyd the data

So Each Table looks something like this, but we need to join all of the tables by the key : Chemical Name, ID, and Region
```{r}
p_value_DMSO_Data
```


## How we #Tidyd the data

So Each Table looks something like this, but we need to join all of the tables by the key : Chemical Name, ID, and Region
Join!: join treatment and Fold Change, and p_val
```{r, results='hide'}
Tidy_DMSO_Data <- left_join(Intensity_DMSO_Data, Fold_Change_DMSO_Data)
            
Tidy_DMSO_Data <- left_join(Tidy_DMSO_Data, p_value_DMSO_Data)
```

## How we #Tidyd the data

So Each Table looks something like this, but we need to join all of the tables by the key : Chemical Name, ID, and Region
Join!: join treatment and Fold Change, and p_val
```{r}
head(Tidy_DMSO_Data)
```


# Subsets of the Data


## Subsets of the Data
-Our data set was very large, so we took subsets of the data

## Subsets of the Data

```{r}
Chems_in_OneCMetabolism <- 
  Tidy_DMSO_Data %>% 
  filter(Chemical_Name== "Cystathionine" |
        Chemical_Name== "DL-Cystathionine" |
        Chemical_Name== "S-Adenosyl-homocysteine" |
        Chemical_Name== "S-Adenosyl-methionine" |
        Chemical_Name== "S-Lactoylglutathione" |
        Chemical_Name== "Glutathione-oxidized" |
        Chemical_Name== "Glutathione-reduceD-")
Chems_in_OneCMetabolism
```


## Subsets of the Data

```{r}
Chems_in_Nucleotides <- 
  Tidy_DMSO_Data %>% 
  filter(Chemical_Name== "5-Methylcytidine" |
        Chemical_Name== "5-Methylcytosine" |
        Chemical_Name== "7-Methylguanine" |
        Chemical_Name== "7-Methylguanosine" |
        Chemical_Name== "Adenine" |
        Chemical_Name== "Adenosine" |
        Chemical_Name== "Adenosine-3-monophosphate" |
        Chemical_Name== "beta-Nicotinamide-adenine-dinucleotide" |
        Chemical_Name== "Guanine" |
        Chemical_Name== "Guanosine" |
        Chemical_Name== "Guanosine-5-tetraphosphate" |
        Chemical_Name== "Cytidine" |
        Chemical_Name== "Cytidine-5-diphosphate-ethanolamine" |
        Chemical_Name== "METHYLGUANIDINE" |
        Chemical_Name== "2-Methylguanosine" |
        Chemical_Name== "Nicotinamide-riboside-cation" |
        Chemical_Name== "Thymine" |
        Chemical_Name== "UMP" |
        Chemical_Name== "3-Methylcytidine" |
        Chemical_Name== "2-Deoxycytidine" |
        Chemical_Name== "2-O-Methylcytidine" |
        Chemical_Name== "3-O-Methylcytidine" |
        Chemical_Name== "Uracil" |
        Chemical_Name== "Uridine" |
        Chemical_Name== "1-Methyladenosine" |
        Chemical_Name== "Xanthine" |
        Chemical_Name== "Xanthosine-5-monophosphate" |
        Chemical_Name== "Inosine" |
        Chemical_Name== "Inosine-5-monophosphate" |
        Chemical_Name== "5-Methylthioadenosine" |
        Chemical_Name== "5-S-Methyl-5-thioadenosine" |
        Chemical_Name== "Uridine-5-diphosphoacetylglucosamine" )
Chems_in_Nucleotides
```


## Subsets of the Data
This is just metabolites with significant p-values and fold changes.
Significant p-val is 0.05, significant fold change = 2

```{r}
DMSO_Data_Sig_Foldchange_AND_pval <- Tidy_DMSO_Data %>% 
  select(Chemical_Name, 
         Region, 
         Fold_Change_vs_Control, 
         p_value_vs_Control, 
         Mean_Intensity, 
         Treatment) %>% 
  filter(Fold_Change_vs_Control >= 2.0 & p_value_vs_Control <= 0.05)
```


## Visualization
```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
DMSO_Data_Sig_Foldchange_AND_pval%>% 
  ggplot(aes(x= Region, fill=Region))+
  geom_bar()+
  theme_dark()+
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust=1))+
  labs(title = "Number of Significant Metabolites by Region ",
       x = "Region of Brain",
       y = "Number of Metabolites")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_brewer(palette = "PuRd")
```

## Visualization
```{r echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
DMSO_Data_Sig_Foldchange_AND_pval%>% 
  ggplot(aes(x= Region, fill=Region))+
  geom_bar()+
  theme_dark()+
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust=1))+
  labs(title = "Number of Significant Metabolites by Region ",
       x = "Region of Brain",
       y = "Number of Metabolites")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_brewer(palette = "PuRd")
```


## Visualization

Looking at the relationship between SAM and SAH is a common indicator on where the equalibrium lays and is idiciative of the activity of the metabolomic pathway
```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
Chems_in_OneCMetabolism %>%
    filter(Chemical_Name== "S-Adenosyl-homocysteine" |
           Chemical_Name== "S-Adenosyl-methionine") %>% 
    ggplot(aes(x = Treatment, y = Mean_Intensity, fill = Region))+
    geom_bar(stat = "identity")+
    theme(axis.text.x = element_text(angle = 60, hjust=1))+
    theme_dark()+
    scale_fill_brewer(palette="PuRd")+
    facet_grid(~Chemical_Name)
```

## Visualization

Looking at the relationship between SAM and SAH is a common indicator on where the equalibrium lays and is idiciative of the activity of the metabolomic pathway
```{r echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
Chems_in_OneCMetabolism %>%
    filter(Chemical_Name== "S-Adenosyl-homocysteine" |
           Chemical_Name== "S-Adenosyl-methionine") %>% 
    ggplot(aes(x = Treatment, y = Mean_Intensity, fill = Region))+
    geom_bar(stat = "identity")+
    theme(axis.text.x = element_text(angle = 60, hjust=1))+
    theme_dark()+
    scale_fill_brewer(palette="PuRd")+
    facet_grid(~Chemical_Name)
```

## Visualization
```{r echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
Chems_in_OneCMetabolism %>%
  filter(Chemical_Name== "Cystathionine" |
        Chemical_Name== "DL-Cystathionine" |
        Chemical_Name== "S-Adenosyl-homocysteine" |
        Chemical_Name== "S-Adenosyl-methionine" |
        Chemical_Name== "Glutathione-oxidized" |
        Chemical_Name== "Glutathione-reduceD-") %>% 
    ggplot(aes(x = Treatment, y = Mean_Intensity, fill = Region))+
    geom_bar(stat = "identity")+
    theme_dark()+
    scale_fill_brewer(palette="PuRd")+
    facet_wrap(~Chemical_Name)+
    theme(axis.text.x = element_text(angle = 60, hjust=1),
    plot.title = element_text(hjust = 0.5))+
    labs(title = "Mean Intensity of Metabolite Signal ",
        x = "Treatment", 
        y = "Mean Intensity")
```













## Visualization

```{r echo=TRUE}
Chems_in_OneCMetabolism %>% 
  ggplot(aes(x=Treatment, y=Chemical_Name))+
  theme_classic(base_family = "Avenir")+
  geom_tile(aes(fill= Mean_Intensity), color="grey")+
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust=1))+
  coord_fixed()+ #maintains aspect ratio so 1 unit on x is 1 unit on y
  labs(title = "Variation in One Carbon Metabolites by Treatment ",
       x = "Treatment",
       y = "Metabolite")+
  theme(plot.title = element_text(hjust = 0.5))
```

## Visualization
```{r echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
Chems_in_OneCMetabolism %>% 
  ggplot(aes(x=Treatment, y=Chemical_Name))+
  theme_classic(base_family = "Avenir")+
  geom_tile(aes(fill= Mean_Intensity), color="grey")+
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust=1))+
  coord_fixed()+ #maintains aspect ratio so 1 unit on x is 1 unit on y
  labs(title = "Variation in One Carbon Metabolites by Treatment ",
       x = "Treatment",
       y = "Metabolite")+
  theme(plot.title = element_text(hjust = 0.5))
```


##Visualization
```{r echo=TRUE}
Chems_in_Nucleotides %>% 
ggplot(aes(x=Treatment, y=Chemical_Name))+
  geom_tile(aes(fill=Mean_Intensity), color="white")+
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 60, hjust=1), 
        plot.title = element_text(hjust = 0.5))+
  coord_fixed()+ #maintains aspect ratio so 1 unit on x is 1 unit on y
  labs(title = "Variation in Nucleotide Metabolites by Treatment ",
       x = "Treatment",
       y = "Metabolite")+
scale_fill_gradientn(colours = terrain.colors(19))
```

##Visualization
```{r echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
Chems_in_Nucleotides %>% 
ggplot(aes(x=Treatment, y=Chemical_Name))+
  geom_tile(aes(fill=Mean_Intensity), color="white")+
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 60, hjust=1), 
        plot.title = element_text(hjust = 0.5))+
  coord_fixed()+ #maintains aspect ratio so 1 unit on x is 1 unit on y
  labs(title = "Variation in Nucleotide Metabolites by Treatment ",
       x = "Treatment",
       y = "Metabolite")+
scale_fill_gradientn(colours = terrain.colors(19))
```

# Thank you!