---
title: "Project"
author: "Zachary Rabow"
date: "2/28/2019"
output:
  pdf_document: default
  html_document: default
---
Load Libraries
```{r}
library("tidyverse")
library("ggplot2")
library("ggthemr")
library("devtools")
library("RColorBrewer")
```

```{r}
options(scipen = 999)
```

Load Data & Rename Columns 
```{r}
Untidy_DMSO_Data <- readr::read_csv("UntidyFYSDataSet.csv")
Untidy_DMSO_Data <- Untidy_DMSO_Data %>%
  rename(
          "Chemical_Name" = ChemicalName,
          "Mean_Intensity_Control" = mean.Control.PBS,
          "Mean_Intensity_02" = mean.0.2.mg.kg.DMSO,
          "Mean_Intensity_20" = mean.2.0.mg.kg.DMSO,
          "Mean_Intensity_40" = mean.4.0.mg.kg.DMSO,
          "Fold_Change_02" = FC_lowvsCTRL,
          "Fold_Change_20" = FC_medvsCTRL,
          "Fold_Change_40" = FC_highvsCTRL,
          "p_val_02" = post.hoc.Control.PBS.0.2.mg.kg.DMSO,
          "p_val_20" = post.hoc.Control.PBS.2.0.mg.kg.DMSO,
          "p_val_40" = post.hoc.Control.PBS.4.0.mg.kg.DMSO
          )
Original_DMSO_Data <- Untidy_DMSO_Data
```

Gather Intensities and make data frame with only intensities
```{r}
Intensity_DMSO_Data<- Untidy_DMSO_Data %>% 
   #Gathers 4 columns 
   gather(Mean_Intensity_Control, 
         Mean_Intensity_02, 
         Mean_Intensity_20, 
         Mean_Intensity_40, 
         key="Treatment", value="Mean_Intensity")%>% 
  
   #deletes part of unwanted text in data
   separate(Treatment, into= c("Delete", "Treatment"), sep = "_Intensity_") %>%   
   # selects 3 "key" columns, treatment, and intensity
   select(Region, ID, Chemical_Name, Treatment, Mean_Intensity)
```

Gather Fold Changes and make data frame with only Fold Changes
```{r}
Fold_Change_DMSO_Data <- Untidy_DMSO_Data %>% 
   
   #Gathers 3 columns 
   gather(Fold_Change_02, 
           Fold_Change_20,
           Fold_Change_40,
          key="Treatment", value="Fold_Change_vs_Control") %>% 
  
   #deletes part of unwanted text in data
   separate(Treatment, into= c("Delete", "Treatment"), sep = "_Change_") %>%   
   # selects 3 "key" columns, treatment, and intensity
   select(Region, ID, Chemical_Name, Treatment, Fold_Change_vs_Control)
```

Gather P-values and make data frame with only P-Values
```{r}
p_value_DMSO_Data<- Untidy_DMSO_Data %>% 
   
   #Gathers 3 columns 
   gather(p_val_02, 
          p_val_20, 
          p_val_40,
          key="Treatment", value="p_value_vs_Control") %>% 
  
   #deletes part of unwanted text in data
   separate(Treatment, into= c("Delete", "Treatment"), sep = "_val_") %>%   
   # selects 3 "key" columns, treatment, and intensity
   select(Region, ID, Chemical_Name, Treatment, p_value_vs_Control)
```

Join!: join treatment and Fold Change, and p_val
```{r}
Tidy_DMSO_Data <- left_join(Intensity_DMSO_Data, Fold_Change_DMSO_Data)
            
Tidy_DMSO_Data <- left_join(Tidy_DMSO_Data, p_value_DMSO_Data)
```

Subsets:

```{r}
Chems_in_OneCMetabolism <- 
  Tidy_DMSO_Data %>% 
  filter(Chemical_Name== "Cystathionine" |
        Chemical_Name== "DL-Cystathionine" |
        Chemical_Name== "S-Adenosyl-homocysteine" |
        Chemical_Name== "S-Adenosyl-methionine" |
        Chemical_Name== "S-Lactoylglutathione" |
        Chemical_Name== "Glutathione-oxidized" |
        Chemical_Name== "Glutathione-reduceD-")
Chems_in_OneCMetabolism
```

```{r}
Chems_in_Nucleotides <- 
  Tidy_DMSO_Data %>% 
  filter(Chemical_Name== "5-Methylcytidine" |
        Chemical_Name== "5-Methylcytosine" |
        Chemical_Name== "7-Methylguanine" |
        Chemical_Name== "7-Methylguanosine" |
        Chemical_Name== "Adenine" |
        Chemical_Name== "Adenosine" |
        Chemical_Name== "Adenosine-3-monophosphate" |
        Chemical_Name== "beta-Nicotinamide-adenine-dinucleotide" |
        Chemical_Name== "Guanine" |
        Chemical_Name== "Guanosine" |
        Chemical_Name== "Guanosine-5-tetraphosphate" |
        Chemical_Name== "Cytidine" |
        Chemical_Name== "Cytidine-5-diphosphate-ethanolamine" |
        Chemical_Name== "METHYLGUANIDINE" |
        Chemical_Name== "2-Methylguanosine" |
        Chemical_Name== "Nicotinamide-riboside-cation" |
        Chemical_Name== "Thymine" |
        Chemical_Name== "UMP" |
        Chemical_Name== "3-Methylcytidine" |
        Chemical_Name== "2-Deoxycytidine" |
        Chemical_Name== "2-O-Methylcytidine" |
        Chemical_Name== "3-O-Methylcytidine" |
        Chemical_Name== "Uracil" |
        Chemical_Name== "Uridine" |
        Chemical_Name== "1-Methyladenosine" |
        Chemical_Name== "Xanthine" |
        Chemical_Name== "Xanthosine-5-monophosphate" |
        Chemical_Name== "Inosine" |
        Chemical_Name== "Inosine-5-monophosphate" |
        Chemical_Name== "5-Methylthioadenosine" |
        Chemical_Name== "5-S-Methyl-5-thioadenosine" |
        Chemical_Name== "Uridine-5-diphosphoacetylglucosamine" )
Chems_in_Nucleotides
```

Significant p-val is 0.05, significant fold change = 2
```{r}
DMSO_Data_Sig_Foldchange_AND_pval <- Tidy_DMSO_Data %>% 
  select(Chemical_Name, 
         Region, 
         Fold_Change_vs_Control, 
         p_value_vs_Control, 
         Mean_Intensity, 
         Treatment) %>% 
  filter(Fold_Change_vs_Control >= 2.0 & p_value_vs_Control <= 0.05)
```

Data Exploration: How many compunds have a sig. p-val, at any dose, in each region
```{r}
Tidy_DMSO_Data %>% 
  select(Chemical_Name, 
         Region, 
         p_value_vs_Control) %>% 
  filter(p_value_vs_Control <= 0.05) %>% 
  count(Region)
```

```{r}
Tidy_DMSO_Data %>% 
  select(Chemical_Name, 
         Region, 
         Fold_Change_vs_Control) %>% 
  filter(Fold_Change_vs_Control >= 2.0) %>% 
  count(Region)
```

```{r}
Tidy_DMSO_Data %>% 
  select(Chemical_Name, 
         Region, 
         Fold_Change_vs_Control, 
         p_value_vs_Control) %>% 
  filter(Fold_Change_vs_Control >= 2.0 & p_value_vs_Control <= 0.05) %>% 
  count(Region)
```

Blood Flow Near regions: First by Cortex, Cerebellum, Basal Ganglia, Hippocampus
```{r}
Significant_and_foldchange_Data %>% 
  ggplot(aes(x= Region, fill=Region))+
  geom_bar()+
  theme_dark(base_family = "Avenir")+
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust=1))+
  labs(title = "Number of Significant Metabolites by Region ",
       x = "Region of Brain",
       y = "Number of Metabolites")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_brewer(palette = "PuRd")
```

```{r}
Chems_in_OneCMetabolism %>% 
  ggplot(aes(x=Treatment, y=Chemical_Name))+
  theme_classic(base_family = "Avenir")+
  geom_tile(aes(fill= Mean_Intensity), color="grey")+
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust=1))+
  coord_fixed()+ #maintains aspect ratio so 1 unit on x is 1 unit on y
  labs(title = "Variation in One Carbon Metabolites by Treatment ",
       x = "Treatment",
       y = "Metabolite")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
Nucleotides <- 
  Tidy_DMSO_Data %>% 
  filter(Chemical_Name== "5-Methylcytidine" |
        Chemical_Name== "5-Methylcytosine" |
        Chemical_Name== "7-Methylguanine" |
        Chemical_Name== "7-Methylguanosine" |
        Chemical_Name== "Adenine" |
        Chemical_Name== "Adenosine" |
        Chemical_Name== "Adenosine-3-monophosphate" |
        Chemical_Name== "beta-Nicotinamide-adenine-dinucleotide" |
        Chemical_Name== "Guanine" |
        Chemical_Name== "Guanosine" |
        Chemical_Name== "Guanosine-5-tetraphosphate" |
        Chemical_Name== "Cytidine" |
        Chemical_Name== "Cytidine-5-diphosphate-ethanolamine" |
        Chemical_Name== "METHYLGUANIDINE" |
        Chemical_Name== "2-Methylguanosine" |
        Chemical_Name== "Nicotinamide-riboside-cation" |
        Chemical_Name== "Thymine" |
        Chemical_Name== "UMP" |
        Chemical_Name== "3-Methylcytidine" |
        Chemical_Name== "2-Deoxycytidine" |
        Chemical_Name== "2-O-Methylcytidine" |
        Chemical_Name== "3-O-Methylcytidine" |
        Chemical_Name== "Uracil" |
        Chemical_Name== "Uridine" |
        Chemical_Name== "1-Methyladenosine" |
        Chemical_Name== "Xanthine" |
        Chemical_Name== "Xanthosine-5-monophosphate" |
        Chemical_Name== "Inosine" |
        Chemical_Name== "Inosine-5-monophosphate" |
        Chemical_Name== "5-Methylthioadenosine" |
        Chemical_Name== "5-S-Methyl-5-thioadenosine" |
        Chemical_Name== "Uridine-5-diphosphoacetylglucosamine" )
Nucleotides
```

```{r} 
Chems_in_Nucleotides %>% 
ggplot(aes(x=Treatment, y=Chemical_Name))+
  geom_tile(aes(fill=Mean_Intensity), color="white")+
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 60, hjust=1), 
        plot.title = element_text(hjust = 0.5))+
  coord_fixed()+ #maintains aspect ratio so 1 unit on x is 1 unit on y
  labs(title = "Variation in Nucleotide Metabolites by Treatment ",
       x = "Treatment",
       y = "Metabolite")+
scale_fill_gradientn(colours = terrain.colors(19))
```

```{r}
OneCMetabolism %>% 
  filter( Region == "Cortex" | Region == "Hippocampus") %>% 
  filter(Treatment == "02" | Treatment == "20" | Treatment== "40") %>% 
  ggplot(aes(x=Chemical_Name, y=Fold_Change_vs_Control, fill=Treatment))+
  theme_dark(base_family = "Avenir")+
  theme(legend.position = "bottom",
      axis.text.x = element_text(angle = 45, hjust=1), 
      plot.title = element_text(hjust = 0.5))+
  labs(title = "Fold Change by Region for One Carbon Metabolites ",
        x = "Metabolite", 
        y = "Fold Change")+
  geom_col(position = "dodge") +
  facet_wrap(~Region)+
  coord_flip()+
  scale_fill_brewer(palette = "PuRd")+
  ylim(0,15)
```

```{r}
Chems_in_Nucleotides%>% 
  filter(p_value_vs_Control < 0.005 & Fold_Change_vs_Control > 2.0)%>%
  ggplot(aes(x=Chemical_Name, y=1/p_value_vs_Control, fill=Treatment))+
  geom_col(position = "dodge") +
  facet_wrap(~Region)+
    theme_dark(base_family = "Avenir")+
    scale_fill_brewer(palette="PuRd")+
      theme(axis.text.x = element_text(angle = 45, hjust=1), 
            plot.title = element_text(hjust = 0.5))+
  labs(title = "Fold Change of Sig. Altered Nucleotides by Treatment and Region ",
        x = "Metabolite", 
        y = "(1/p-value)")
```


```{r}
Chems_in_OneCMetabolism %>%
  filter(Chemical_Name== "Cystathionine" |
        Chemical_Name== "DL-Cystathionine" |
        Chemical_Name== "S-Adenosyl-homocysteine" |
        Chemical_Name== "S-Adenosyl-methionine" |
        Chemical_Name== "Glutathione-oxidized" |
        Chemical_Name== "Glutathione-reduceD-") %>% 
    ggplot(aes(x = Treatment, y = Mean_Intensity, fill = Region))+
    geom_bar(stat = "identity")+
    theme_dark()+
    scale_fill_brewer(palette="PuRd")+
    facet_wrap(~Chemical_Name)+
    theme(axis.text.x = element_text(angle = 60, hjust=1),
    plot.title = element_text(hjust = 0.5))+
    labs(title = "Mean Intensity of Metabolite Signal ",
        x = "Treatment", 
        y = "Mean Intensity")
```

Looking at the relationship between SAM and SAH is a common indicator on where the equalibrium lays and is idiciative of the activity of the metabolomic pathway
```{r}
Chems_in_OneCMetabolism %>%
    filter(Chemical_Name== "S-Adenosyl-homocysteine" |
           Chemical_Name== "S-Adenosyl-methionine") %>% 
    ggplot(aes(x = Treatment, y = Mean_Intensity, fill = Region))+
    geom_bar(stat = "identity")+
    theme_dark()+
    scale_fill_brewer(palette="PuRd")+
    facet_grid(~Chemical_Name)+
    theme(axis.text.x = element_text(angle = 45, hjust=1),
          plot.title = element_text(hjust = 0.5))+
    labs(title = "SAM : SAH Intensity  ",
        x = "Treatment", 
        y = "Mean Intensity")
```
